### 2019-01-10 Data_extractions.Rmd
add html file and starting file for the next exercise

---
title: "Data_extractions"
author: "ML"
date: "10 January 2019"
output: html_document
---

```{r setup, cache = F, echo=FALSE}
knitr::opts_chunk$set(error = TRUE) #allow some execution errors for demonstration purposes
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, warning = FALSE)
```

# EXERCISES - PHYLOGENETIC TREES FOR META-ANALYSIS

Prepared by Malgorzata (Losia) Lagisz, BEES, UNSW, AU, losialagisz@yahoo.com.

Introduction  
-------

**TO DO**:

Prepare a set of 3 DAH papers to be extracted using *metaDigitise* package
See: https://github.com/daniel1noble/metaDigitise (link to the paper?)

USE ES_calculator to calculate ES?

Start from installing (as required) and loading R packages.

```{r prepare, message = FALSE}
#install.packages("devtools","fulltxt","ape","metafor")

```