### 2019-02-07 R/Trees_making_solutions.Rmd
reorganised

---
title: "Trees_making_solutions"
author: "ML"
date: "1/16/2019"
output: html_document
---

```{r prepare, message = FALSE, echo = FALSE, eval = TRUE, warning = FALSE}
knitr::opts_chunk$set(error = TRUE) #allow some execution errors for demonstration purposes
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, warning = FALSE) # install.packages("diagram")
# install.packages("tidyverse")
# install.packages("ape")
# install.packages("curl")
# install.packages("fulltext")
# install.packages("treebase")
# install.packages("devtools")
# devtools::install_github("ropensci/rotl", dependencies = TRUE, build_vignette=TRUE)
# install.packages("rotl")
library(diagram, tidyverse)
library(ape, curl)
library(fulltext, metafor)
library(treebase, devtools)
library(rotl)
```

********************************************************************************

### EXAMPLE 5 - try it yourself - rotl

Use *rotl* to create a tree for the bunch of reptiles.   
**Remember**: Make sure that species names match exactly tree tip labels

```{r rotl solution, echo = TRUE, eval = TRUE, warning = FALSE}
myspecies2 <- c("Gekko japonicus", "Geochelone sulcata", "Gopherus agassizii", "Gopherus polyphemus", "Graptemys ouachitensis", "Graptemys pseudogeographica", "Hemidactylus bowringii", "Heterodon platirhinos", "Heteronotia binoei", "Iberolacerta cyeni", "Lacerta schreiberi", "Lampropholis delicata", "Lampropholis guichenoti", "Liasis fuscus")

taxa2 <- tnrs_match_names(names = myspecies2)
taxa2 #visual check

tree2 <- tol_induced_subtree(ott_ids= taxa2[["ott_id"]], label_format = "name") #dont worry about the Warning message
plot(tree2, cex=.8, label.offset =.1, no.margin = TRUE)
tree2$tip.label <- gsub("_"," ", tree2$tip.label) #get rid of underscores
# tree2$tip.label
#tree$node.label <- NULL #delete node labels

#check if the tree2 is really binary 
is.binary.tree(tree2) #TRUE (i.e. there are no polytomies)

#check the overlap between species name from the list nad the tree2 
intersect(as.character(tree2$tip.label), myspecies2) #12 matching names
setdiff(myspecies2, as.character(tree2$tip.label)) #2 mismatches - these are only in myspecies2
setdiff(as.character(tree2$tip.label), myspecies2) #2 mismatches - these are only in the tree2$tip.label

#rename the tip labels on the tree to match the species names from our list
tree2$tip.label <- gsub("Centrochelys sulcata", "Geochelone sulcata", tree2$tip.label) 
tree2$tip.label <- gsub("Iberolacerta cyreni", "Iberolacerta cyeni", tree2$tip.label) 
#Note: you should actually check which name from each pair is "more correct" and accordingly check the names either on the tree or in your species list / data set

#re-check the overlap between species name from the list nad the tree2 
intersect(as.character(tree2$tip.label), myspecies2) #14 matching names
setdiff(myspecies2, as.character(tree2$tip.label)) #0 mismatches
setdiff(as.character(tree2$tip.label), myspecies2) #0 mismatches
```
Done!

********************************************************************************

### EXAMPLE 6 - try it yourself- treebase

Use *treebase* to create a tree for the three Quercus species from dat.Curtis1998.

**Hint**: Quercus prinus is also called Quercus montana (https://en.wikipedia.org/wiki/Quercus_montana).   
Once you substitute, all three species can be found on Tree 113123 of Study 22057, and also in several other studies and trees

```{r treebase solution, echo = TRUE, eval = TRUE, warning = FALSE}
#Code you need intially to get a list of species from Curtis 1998, in case you did not run Example 3:
dat <- metafor::dat.curtis1998
species_Curtis1998 <- paste(dat$genus, dat$species, sep=" ") 
species_Curtis1998 <- unique(tolower(species_Curtis1998)) #make species name unique
species_Curtis1998 <- gsub("populusx euramericana", "populus euramericana", species_Curtis1998) #merge 2 partially overlapping names (typo? or hybrid?) - remember to fix this in the main data set!
species_Curtis1998 <- unique(tolower(species_Curtis1998)) #make the names unique again
species_Curtis1998 <- paste(toupper(substr(species_Curtis1998, 1, 1)), substr(species_Curtis1998, 2, nchar(species_Curtis1998)), sep="") #capitalise first letters

#extract three Quercus species names:
Quercus3sp <- species_Curtis1998[grep("Quercus ", species_Curtis1998)]
Quercus3sp #3 species
Quercus3sp <- gsub(" ","_", Quercus3sp) #put underscores back in the names
Quercus3sp <- gsub("prinus","montana", Quercus3sp) #change the name to its currentky used synonym

Quercus3sp_string <- paste(Quercus3sp, sep = "", collapse = ' and ') 
Quercus3sp_string #you should get "Quercus_rubrum and Quercus_saccharinum and Quercus_pensylvanicum and Quercus_saccharum"

Quercus_treebase <- search_treebase(Quercus3sp_string, by=c("taxon","taxon","taxon"), max_trees=10, exact_match = TRUE, only_metadata = TRUE) #note different commas than in the treebase tutorial!
t(data.frame(matrix(unlist(Quercus_treebase), nrow=6, byrow=F), row.names = c("S.id", "Tr.id", "type", "kind", "quality", "ntax"))) #note the study IDs that you are using - you can find these on Treebase
Quercuss_tree <- search_treebase("113123", by="id.tree", exact_match = TRUE) #returns a single tree in a list 
Quercuss_tree <- Quercuss_tree[[1]] #extract tree from the list
#if Treebase does not work use: Quercuss_tree <- read.nexus(file = "T113123.nex") #read Nexus tree file
intersect(as.character(Quercuss_tree$tip.label), Quercus3sp) # all 3 are present and matching
# plot(Quercuss_tree, cex=.8, label.offset =.1, no.margin = TRUE) #a big tree of all sorts of trees - needs good trimming!
Quercus3sp_tree <- drop.tip(Quercuss_tree, Quercuss_tree$tip.label[-match(Quercus3sp, Quercuss_tree$tip.label)])
#plot(Quercus3sp_tree, cex=.8, label.offset =.1, no.margin = TRUE)

#check the overlap between species name from the data fila nad the tree file
intersect(as.character(Quercus3sp_tree$tip.label), Quercus3sp) # all matching, but we changed the species name in the list, so it will no longer match the original dataset
Quercus3sp_tree$tip.label <- gsub("montana", "prinus", Quercus3sp_tree$tip.label) #replace the tip label back with our species name
Quercus3sp_tree$tip.label <- gsub("_"," ", Quercus3sp_tree$tip.label) #get rid of the underscores
```

```{r plot Quercus3sp_tree, fig.width=10, fig.height=3, echo=TRUE, message=FALSE}
plot(Quercus3sp_tree, cex=.8, label.offset =.1, no.margin = TRUE)
```

We first need to prune given branch from the main tree to remove all but one species affected by polytomy. Then we can replace this species with the resolved sub-tree (Quercus3sp_tree):

```{r graft Quercus_tree}
tree <- read.tree(file="tree_curtis1998.tre") #run if you need to read it in
tree_pruned <- drop.tip(tree, c("Quercus_prinus", "Quercus_alba")) #prune 3 species
tree_grafted <- bind.tree(tree_pruned, Quercus3sp_tree, where=which(tree_pruned$tip.label=="Quercus_rubra")) #graft 3 species to the remaining Acer species
```

```{r plot tree_grafted Acers, fig.width=10, fig.height=8, echo=TRUE, message=FALSE}
plot(tree_grafted, cex=.8, label.offset =.1, no.margin = TRUE)
```

Done!